---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# infun

<!-- badges: start -->
<!-- badges: end -->

This is a collection of R utilities functions for me, but maybe also for you.

Functions may be added, specifications of functions may change or become obsolete, and names may change without notice.

## Installation

install the development version install from GitHub:

```r
install.packages("remotes")
remotes::install_github("indenkun/infun")
```

## Example

load library.

```{r example}
library(infun)
```

Make a sample data for README (`example.data`).

```{r}
example.data <- data.frame(value1 = 1:10,
                           value2 = c(1:3, "strings", 5:10),
                           value3 = c(1:3, "strings", 5:10),
                           value4 = 11:20)
```

### `find.not.numeric.value()`

This function is used to find the where in the vector there are values that cannot be converted to numbers.

If you specify a column from the dataframe with `[]`, it behaves in the same way.

If you input a dataframe that contains multiple columns, it will return the location of the column that contains the value that cannot be converted to a number, if specified.

The fourth data in value2 of `example.data` will be a string. `find.not.numeric.value()` will show where all the data in the vector is located if there is a value that will be forced to NA when converted to numeric type by `as.numeric()`. If there is no value to be converted, NA is returned.

```{r}
find.not.numeric.value(example.data$value1)
find.not.numeric.value(example.data$value2)
find.not.numeric.value(example.data[1])
find.not.numeric.value(example.data[2])
find.not.numeric.value(example.data)
```

### `find.same.value.col()`

This function is used to find a column consisting of the same value in a data frame.

If you run `same.value.col()` on `example.data`, you will see that the second and third columns of the sample data all have the same value.

The result is returned in a list format.

```{r}
find.same.value.col(example.data)
```

`unique_col()` is a function to remove duplicate columns in a data frame, the column version of `{base}`'s `unique()`.

```{r}
unique_col(example.data)
```

### `find.not.integer.value()`

This function is used to find a non-integer value in a vector.

The input value can be of any type, but it must be a vector of numbers only.
If a string or other value is entered, a warning message will be displayed and `NA` will be returned.
If you get a warning message that a non-numeric value is entered, try `find.not.numeric.value()` to find the non-numeric value.

If you input a dataframe that contains multiple columns, it will return the location of the column that contains the value that non-integer, if specified.

```{r}
example.data.integer <- data.frame(Item1 = 1:10,
                                   Item2 = c(1:5, 6.5, 7.5, 8:10),
                                   Item3 = c(1:6, "strings", 8:10))
```

Returns the location as a number if the value is not an integer. If no non-integer values are entered in a vector consisting of numbers, NA will be returned.

If `"logical"` is specified in `where`, a vector of logical type will be returned.

```{r, warning=TRUE}
find.not.integer.value(example.data.integer$Item1)
find.not.integer.value(example.data.integer[2])
find.not.integer.value(example.data.integer[1:2])
find.not.integer.value(example.data.integer$Item3)
```

### `find.not.as.Date.value()`

This function is used to find the where in the vector there are values that cannot be converted to `Date` using `as.Date()` in `{base}`.

```{r}
example.data.Date <- data.frame(Date1 = c("2021-7-28", "2021-08-08", "2021-08-24", "2021-09-05"),
                                Date2 = c("2021-7-28", "NOTDATE", "NOTDATE", "2021-09-05"),
                                Date3 = c("210728", "21/08/08", "21/Aug-24", "21sep5"))

find.not.as.Date.value(example.data.Date$Date1)
find.not.as.Date.value(example.data.Date$Date2)
find.not.as.Date.value(example.data.Date$Date3)
```

### `find.not.as_date.value()`

This function is used to find the where in the vector there are values that cannot be converted to `Date` using `as_date()` in `{lubridate}`.

There is a slight difference between the values that can be converted to Date by `{lubridate}`'s `as_date()` and those that can be converted by `{base}`'s `as.Date()`.

```{r}
find.not.as_date.value(example.data.Date$Date1)
find.not.as_date.value(example.data.Date$Date2)
# as_date() converts even relatively fuzzy forms if they can be changed to a date type, while as.Date() operates relatively more strictly.
find.not.as_date.value(example.data.Date$Date3)
```

### `add.str()`

Combine all the items in a specific column of a data frame with any string of characters in the original data frame. The converted column will be a string because it contains strings such as ALL.

You need to specify any column as `key` with the column name.

```{r}
example.data.add.all <- add.str(example.data, "value1")
head(example.data.add.all, 20)
```

### `random.Date()`

`random.Date()` is a function that randomly creates a vector of dates at a specified sample size between a specified date and a date.

```r
random.Date(from = "2021/1/1", to = "2021/4/1", size = 10)
#>  [1] "2021-01-21" "2021-03-17" "2021-03-09" "2021-02-20" "2021-02-24"
#>  [6] "2021-02-22" "2021-02-06" "2021-03-24" "2021-02-03" "2021-03-11"
```

### `age.cal()`

`age.cal()` is a function that calculates the number of years (age by default), months, and days from a specified date to a specified date.

```{r}
age.cal(from = c("2000/1/1", "2010/1/1"), to = "2021/4/1")
```


### `tableone.rename.*()`

These functions are used to change the headline character in the item name of a table created with `{tableone}`'s to any character.

`tableone.rename.overall()` is used to change the "Overall" character in the item name of a table created with `{tableone}`'s `CreateTableOne()` to any character.

```{r}
# This is the code to create a sample table in `{tableone}`.
library(tableone)

iris.table <- CreateTableOne(data = iris)
iris.table

# Rename "Overall" to "ALL".
tableone.rename.overall(iris.table, rename.str = "ALL")
```

`tableone.rename.headline()` is a function that change any heading (including Overall) to any character by setting the table heading as an formula before and after the change.

```{r}
# This is the code to create a sample table in `{tableone}`.
library(tableone)
library(survival)
data(pbc)

varsToFactor <- c("status","trt","ascites","hepato","spiders","edema","stage")
pbc[varsToFactor] <- lapply(pbc[varsToFactor], factor)
vars <- c("time","status","age","sex","ascites","hepato",
          "spiders","edema","bili","chol","albumin",
          "copper","alk.phos","ast","trig","platelet",
          "protime","stage")
tableOne <- CreateTableOne(vars = vars, strata = c("trt"), data = pbc, addOverall = TRUE)
tableOne

# Rename headline name "1" to "D-penicillmain", "2" to "placebo".
# Names that contain hyphens will be evaluated as negative in the formula, so they must be enclosed in quotation marks.
tableone.rename.headline(tableOne, rename.headline = list(1 ~ "D-penicillmain", 2 ~ placebo))

```

### `seq_geometric()`

This function is used to generate a sequence of equal ratios, also known as a geometric sequence.

By specifying the first term in `from`, the last term or the closest value to the last term in `to`, and the common ratio in `by.rate`, you can obtain an geometric sequence of "first term * common ratio ^ n" from "from" to the closest value to "to".

```{r}
seq_geometric(from = 1, to = 128, by.ratio = 2)
```

### `Rtools.pacman.package.*()`

These are functions to search for packages that can be installed by Rtools' pacman. In short, it is a wrapper for some of the functions of pacman in Rtools.

Cannot be used except in a Windows environment where Rtools40 or later is installed.

`Rtools.pacman.package.list()` is a function that outputs a list of packages that can be installed by Rtools pacman from repository. By specifying arguments, you can extract only those packages that are already installed, or only those that are yet uninstalled.

```{r}
package.list <- Rtools.pacman.package.list()
# It's too long, so show part of it in head()
head(package.list)
```

`Rtools.pacman.package.list()` is a function that displays a list of packages that can be installed by pacman in Rtools from repository with the specified arguments in the string. If no matching package is found, return NA.

```{r}
Rtools.pacman.package.find("curl")
```

### `scale.data.frame()`

`scale.data.frame()` is generic function whose default method centers and/or scales the columns of a numeric in data frame.
The non-numeric values in the data frame will remain unchanged.

In short, it is a generic function of `{base}` `cale()`.

It is a generic function of `scale()`, so call it with `scale()` when `{infun}` library is loaded. If the object is a data frame, this will work by itself.

If you want to call it explicitly, use `infun:::scale.data.frame()`.

If you want to explicitly use the `{base}` `scale()` after loading `{infun}` as a library, you can use it in `scale.default()`.

```{r}
z.iris <- scale(iris)
# It's too long, so show part of it in head()
head(z.iris)
```

### `save_gtsummary()`

This function is used to output the table created by the gtsummary package in PowerPoint or Word, or as an image file.
It just wraps `{gtsummary}`'s `as_flex_table()` and `{flextalbe}`'s `save_as*()` functions.

Supported filename extensions: .pptx, .docx, .png, .pdf, .jpg.

```r
library(gtsummary)
library(tidyverse)

# Sample code for gtsummary
tbl_summary_ex1 <-
  trial %>%
  select(age, grade, response) %>%
  tbl_summary()
  
# Output the table created by gtsummary to PowerPoint(.pptx).
tbl_summary_ex1 %>% 
  save_gtsummary(path = "table.pptx")
```
### `round_any_*()`

`round_any()` is used to round a vector made of numbers to an approximation of a sequence of numbers with arbitrary equidifferences.

If the value matches an arbitrary isoperimetric sequence, the value will be output as is.

If the `type` argument is `ceiling`, it will round to the upper side of the nearest value, and if the `type` argument is `floor`, it will round to the lower side.

```{r}
example.vector <- seq(0, 1, 0.1)
example.vector

round_any(example.vector, by = 0.25, type = "ceiling")

round_any(example.vector, by = 0.25, type = "floor")
```

`round_any_ceiling()` is a simplified version of `round_any()`, which outputs the result with the argument of `type` fixed to `ceiling` and `origin` fixed to `0`.
`round_any_floor()` is a simplified version of `round_any()`, where the `type` argument is fixed to `floor` and the `origin` is fixed to `0`.
`round_any_*` is faster than `round_any()` in most cases, because the internal processing is done as a vector.

**However, in rare cases, `round_any_*()` may not be possible to obtain accurate values because of R's internal floating point arithmetic. `round_any()` creates a sequence of numbers and compares them, so it gives accurate rounding results.**

```{r}
round_any_ceiling(example.vector, 0.25)

round_any_floor(example.vector, 0.25)
```

### `rand_moji()`

Function to create a random Japanese (Kanji or Hiragana) string. 
Only the range of regular kanji is supported.

It is also compatible and reproducible for `set.seed()`.

``` r
rand_moji(length = 3, size = 3, moji = "kanji")
#> [1] "缶販微" "症凸噴" "侍沖忌"

rand_moji(length = 3, size = 3, moji = "hiragana")
#> [1] "にうぁ" "もんゐ" "えヴり"
```

It is a random string, so it does not reflect the normal rules of Japanese. 
In the case of hiragana, characters that do not normally appear at the beginning of a string, such as Sutegana and "n", will also appear at the beginning.

Katakana strings are not supported and should be converted using functions such as `stringi::stri_trans_general()` in the `{stringi}` package.

``` r
hiragana.moji <- rand_moji(length = 3, size = 3, moji = "hiragana")
hiragana.moji
#> [1] "つざそ" "がせど" "せこへ"
katakana.moji <- stringi::stri_trans_general(hiragana.moji, "hiragana-katakana")
katakana.moji
#> [1] "ツザソ" "ガセド" "セコヘ"
```
### `str_remove_sandwich()`

Delete a string of characters sandwiched between specific characters.

The specified string must be a single character, and the first and last characters of the string must be different.

```{r}
str_remove_sandwich("西馬音内《にしもない》は雄勝郡羽後町《おがちぐんうごまち》です。", start_pattern = "《", end_pattern = "》")
```

Please escape characters that need to be escaped, such as `()`.

```{r}
str_remove_sandwich("dplyr (≥ 0.8.3), arabic2kansuji (≥ 0.1.0)", "\\(", "\\)")
```

### `subset_interchange_col()`

For any two columns specified in the data frame (say column A and B), if the combination of column A and B is the same even if they are swapped, it will return it as a data frame or a row number.

For example, if column A has "TOM" and "BOB", and the same respective row in column B has "BOB" and "TOM", the row will be extracted as interchangeable.

Also, when there is a row with the same value in column A and B, it is also determined to be interchangeable and extracted.

```{r}
example.interchange <- data.frame(X = c("TOM", "BOB", "JOHN", "POP"),
                                  Y = c("BOB", "TOM", "BEE", "TOO"),
                                  Z = seq(10, 40, by = 10))
subset_interchange_col(example.interchange, "X", "Y")
subset_interchange_col(example.interchange, "X", "Y", out.put = "num")
```

### `list2data.frame_*()`

Function to convert a list into a dataframe.

For list of different lengths, the data frame is constructed according to the longest list, and for short lists, the missing places are filled with NA according to the long list.

`list2data.frame_cbind()` makes each element of the list a column.

`list2data.frame_rbind()` makes each element of the list a row.

```{r}
multi_length_list <- list(A = 1,
                          B = 1:2,
                          C = 1:3,
                          D = c(1, NA, 3:4),
                          E = c(1, NA))
list2data.frame_cbind(multi_length_list)
list2data.frame_rbind(multi_length_list)
```

Of course, lists of the same length can also be converted to data frames.

```{r}
equal_length_list <- list(a = 1:4,
                          b = 5:8,
                          c = 9:12)
list2data.frame_cbind(equal_length_list)
list2data.frame_rbind(equal_length_list)
```


## Imports packages

* `{purrr}`

## Suggests packages

* `{gtsummary}`
* `{flextable}`
* `{tools}`
* `{lubridate}`
* `{dplyr}`

## License

MIT.

## Notice

* The email address listed in the DESCRIPTION is a dummy. If you have any questions, please post them on ISSUE.
