---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# infun

<!-- badges: start -->
<!-- badges: end -->

This is a collection of R utilities functions for me, but maybe also for you.

Functions may be added, specifications of functions may change or become obsolete, and names may change without notice.

## Installation

install the development version install from GitHub:

```r
install.packages("remotes")
remotes::install_github("indenkun/infun")
```

## Example

load library.

```{r example}
library(infun)
```

Make a sample data for README (`example.data`).

```{r}
example.data <- data.frame(value1 = 1:10,
                           value2 = c(1:3, "strings", 5:10),
                           value3 = c(1:3, "strings", 5:10),
                           value4 = 11:20)
```

### `find.not.numeric.value()`

This function is used to find the where in the vector there are values that cannot be converted to numbers.

The fourth data in value2 of `example.data` will be a string. `find.not.numeric.value()` will show where all the data in the vector is located if there is a value that will be forced to NA when converted to numeric type by `as.numeric()`. If there is no value to be converted, NA is returned.

```{r}
find.not.numeric.value(example.data$value1)
find.not.numeric.value(example.data$value2)
```

### `find.same.value.col()`

This function is used to find a column consisting of the same value in a data frame.

If you run `same.value.col()` on `example.data`, you will see that the second and third columns of the sample data all have the same value.

The result is returned in a list format.

```{r}
find.same.value.col(example.data)
```

### `add.str()`

Combine all the items in a specific column of a data frame with any string of characters in the original data frame. The converted column will be a string because it contains strings such as ALL.

You need to specify any column as `key` with the column name.

```{r}
example.data.add.all <- add.str(example.data, "value1")
head(example.data.add.all, 20)
```

### `random.Date()`

`random.Date()` is a function that randomly creates a vector of dates at a specified sample size between a specified date and a date.

```r
random.Date(from = "2021/1/1", to = "2021/4/1", size = 10)
#>  [1] "2021-01-21" "2021-03-17" "2021-03-09" "2021-02-20" "2021-02-24"
#>  [6] "2021-02-22" "2021-02-06" "2021-03-24" "2021-02-03" "2021-03-11"
```

### `age.cal()`

`age.cal()` is a function that calculates the number of years (age by default), months, and days from a specified date to a specified date.

```{r}
age.cal(from = c("2000/1/1", "2010/1/1"), to = "2021/4/1")
```


### `tableone.overall.rename()`

This function is used to change the "Overall" character in the item name of a table created with `{tableone}`'s `CreateTableOne()` to any character.

```{r}
library(tableone)

iris.table <- CreateTableOne(data = iris)
iris.table

# Rename "Overall" to "ALL".
tableone.overall.rename(iris.table, rename.str = "ALL")
```

## `seq_geometric()`

This function is used to generate a sequence of equal ratios, also known as a geometric sequence.

By specifying the first term in `from`, the last term or the closest value to the last term in `to`, and the common ratio in `by.rate`, you can obtain an geometric sequence of "first term * common ratio ^ n" from "from" to the closest value to "to".

```{r}
seq_geometric(from = 1, to = 128, by.ratio = 2)
```

## `Rtools.pacman.package.*()`

These are functions to search for packages that can be installed by Rtools' pacman. In short, it is a wrapper for some of the functions of pacman in Rtools.

Cannot be used except in a Windows environment where Rtools40 or later is installed.

`Rtools.pacman.package.list()` is a function that outputs a list of packages that can be installed by Rtools pacman from repository. By specifying arguments, you can extract only those packages that are already installed, or only those that are yet uninstalled.

```{r}
package.list <- Rtools.pacman.package.list()
# It's too long, so show part of it in head()
head(package.list)
```

`Rtools.pacman.package.list()` is a function that displays a list of packages that can be installed by pacman in Rtools from repository with the specified arguments in the string. If no matching package is found, return NA.

```{r}
Rtools.pacman.package.find("curl")
```

## `scale.data.frame()`

`scale.data.frame()` is generic function whose default method centers and/or scales the columns of a numeric in data frame.
The non-numeric values in the data frame will remain unchanged.

In short, it is a generic function of `{base}` `cale()`.

It is a generic function of `scale()`, so call it with scale when `{infun}` library is loaded. If the object is a data frame, this will work by itself.

If you want to call it explicitly, use `infun:::scale.data.frame()`.

```{r}
z.iris <- scale(iris)
# It's too long, so show part of it in head()
head(z.iris)
```


## Imports packages

* `{purrr}`

## Suggests packages

* `{tableone}`

## License

MIT.
